---
title: "TITANIC MACHINE LEARNING REPORTING"
author: "Daouda Tandiang DJIBA"
output: github_document
---

#Data Analysis from Titanic disaster

```{r, echo=FALSE}
setwd("C:\\Users\\dell\\Desktop\\DATA SCIENCE PROJECT\\TITANIC MACHINE LEARNING FROM DISASTER")
train <- read.csv("train.csv", header=TRUE)
test <- read.csv("test.csv", header=TRUE)

```

##Description of the data:

Train dataset:
Variables Names, types and summary on the train dataset: 
```{r, echo=FALSE}
names(train)
str(train)
summary(train)
```

Test dataset:
Variables Names, types and summary on the train dataset:
```{r, echo=FALSE}
names(test)
str(test)
summary(test)
```
To avoid duplicate data work, let's combine the train dataset and the test  dataset to a dataframe called data.combined. 

To do this, let's first add the "survived" variable to the Test dataset.

```{r}
test.survived <- data.frame(survived = rep("None", nrow(test)), test[,])
str(test.survived)
```
Combine data sets
```{r}
data.combined <- rbind(train, test.survived)
str(data.combined)
```
Let's Transform the "Pclass" and the "Survived" variables to factor for analysis:
```{r =eval(1,2)}
data.combined$survived <- as.factor(data.combined$survived)
data.combined$pclass <- as.factor(data.combined$pclass)
str(data.combined)
```
##Workforce distribution for some variables

- "Survived" variable:

Proportions for the first 891 rows (Train dataset)
```{r}
prop.table(table(data.combined$survived [1:891])) * 100
barplot(prop.table(table(data.combined$survived [1:891])) * 100, ylim=c(0,50), las=1 )
```

61.61 % passengers perished

38.38 % passengers survived

- "Pclass" variable

Distribution across classes for the first 891 rows (Train dataset)
```{r}
table(data.combined$pclass[1:891])
barplot(table(data.combined$pclass[1:891]), ylim=c(0, 500),las=1)
```

216 passengers in the first class

184 passengers in the second class

491 passengers in the third class


Let's take a look to the survival rate distribution in classes

###Hypothesis: Rich passengers survived at a higher rate

Distrution of the "survived" variable in classes

```{r}
library(ggplot2)
library(scales)  # for percentage scales
ggplot(data.combined[1:891,], aes(x = pclass, fill = factor(survived))) +
  geom_bar() +
  xlab("Pclass") +
  ylab("Total Count") +
  labs(fill = "Survived") 
```

```{r}
ggplot(data.combined[1:891,], aes(x = pclass, fill = survived )) +
geom_bar(aes(y = ..count..))+
geom_text(aes(label = ..count.., y = ..count..), stat="count", position = position_stack(vjust = 0.5)) +
xlab("Pclass") + 
ylab("Total Count") + 
labs(fill = "survived") +
facet_grid(~survived)
```
