---
title: "TITANIC MACHINE LEARNING REPORTING"
author: "Daouda Tandiang DJIBA"
output: github_document
---

#Data Analysis from Titanic disaster

```{r, echo=FALSE}
setwd("C:\\Users\\dell\\Desktop\\DATA SCIENCE PROJECT\\TITANIC MACHINE LEARNING FROM DISASTER")
train <- read.csv("train.csv", header=TRUE)
test <- read.csv("test.csv", header=TRUE)

```

##Description of the data:

Train dataset:
Variables Names, types and summary on the train dataset: 
```{r, echo=FALSE}
names(train)
str(train)
summary(train)
```

Test dataset:
Variables Names, types and summary on the train dataset:
```{r, echo=FALSE}
names(test)
str(test)
summary(test)
```
To avoid duplicate data work, let's combine the train dataset and the test  dataset to a dataframe called data.combined. 

To do this, let's first add the "survived" variable to the Test dataset.

```{r}
test.survived <- data.frame(survived = rep("None", nrow(test)), test[,])
str(test.survived)
```
Combine data sets
```{r}
data.combined <- rbind(train, test.survived)
str(data.combined)
```
Let's Transform the "Pclass" and the "Survived" variables to factor for analysis:
```{r =eval(1,2)}
data.combined$survived <- as.factor(data.combined$survived)
data.combined$pclass <- as.factor(data.combined$pclass)
str(data.combined)
```
##Workforce distribution for some variables

- "Survived" variable:

Proportions for the first 891 rows (Train dataset)
```{r}
prop.table(table(data.combined$survived [1:891])) * 100
barplot(prop.table(table(data.combined$survived [1:891])) * 100, ylim=c(0,50), las=1 )
```

61.61 % passengers perished

38.38 % passengers survived

- "Pclass" variable

Distribution across classes for the first 891 rows (Train dataset)
```{r}
table(data.combined$pclass[1:891])
barplot(table(data.combined$pclass[1:891]), ylim=c(0, 500),las=1)
```

216 passengers in the first class

184 passengers in the second class

491 passengers in the third class


Let's take a look to the survival rate distribution in classes

###Hypothesis: Rich passengers survived at a higher rate

Distrution of the "survived" variable in classes

```{r}
library(ggplot2)
library(scales)  # for percentage scales
ggplot(data.combined[1:891,], aes(x = pclass, fill = factor(survived))) +
  geom_bar() +
  xlab("Pclass") +
  ylab("Total Count") +
  labs(fill = "Survived") 
```

```{r}
ggplot(data.combined[1:891,], aes(x = pclass, fill = survived )) +
geom_bar(aes(y = ..count..))+
geom_text(aes(label = ..count.., y = ..count..), stat="count", position = position_stack(vjust = 0.5)) +
xlab("Pclass") + 
ylab("Total Count") + 
labs(fill = "survived") +
facet_grid(~survived)
```

Persons in the first class have more chance to survived than perishing.

The survival rate in the first class is better than the survival rate in the second and the third class.

From data analysis perpective, it seems that Pclass is a important variable that can explain who survived and who did not.

###Let's take a look on the "name" variable

Examine the first few names in the training data set

```{r, echo=FALSE}
head(as.character(train$name))
```

How many unique names are in the train set and the test set?

```{r}
length(unique(as.character(data.combined$name)))
```

1 037 names are unique and 2 duplicated names:

```{r}
dup.names <- as.character(data.combined[which(duplicated(as.character(data.combined$name))), "name"])
data.combined[which(data.combined$name %in% dup.names),]
```

These names are common in England. It represent different persons

We also see that names are formated in a specific way with a Title.

Let's take a look closer to the Title name

Create a utility function to help with title extraction
NOTE - Using the grep function here

```{r}
extractTitle <- function(name) {
  name <- as.character(name)
  if(length(grep("Miss.", name)) > 0) {
    return("Miss.")
  } else if(length(grep("Master.", name)) > 0) {
    return("Master")
  } else if(length(grep("Mrs.", name)) > 0) {
    return("Mrs")
  } else if(length(grep("Mr.", name)) > 0) {
    return("Mr")
  } else {
    return("other")
  }
    
}

titles <- NULL 
for(i in 1:nrow(data.combined)){
  titles <- c(titles, extractTitle(data.combined[i, "name"]))
  
}
data.combined$title <- as.factor(titles)
```

Let's plot the relationship between Pclass, Survived and Title variables.

```{r}
ggplot(data.combined[1:891,], aes(x = title, fill = survived))+
  geom_bar() +
  facet_wrap(~pclass) +
  ggtitle("Pclass") + 
  xlab("Title") +
  ylab("Total Count") +
  labs(fill = "Survived")
```
